---
title: "World Bank - Income Distributions and GDP"
author: "mcdevitt"
date: "8 mars 2017"
output:
  md_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(repmis)
library(RCurl)
library(methods)
library(ggplot2)
library(pastecs)

require (reshape2)
require (xtable)
require (ascii)

home_dir <- "~/_smu/msds 6306 - data science/case_study_1/gdp_worldbank"

data_dir <- paste0(home_dir, "/data")
analysis_dir <- paste0(home_dir, "/analysis")
summary_dir <- paste0(home_dir, "/summary")
```


## R Markdown

####1 : Merge the data based on the country shortcode. How many of the IDs match?  


####2 : Sort the data frame in ascending order by GDP (so United States is last). What is the 13th country in the resulting data frame?  


####3 : What are the average GDP rankings for the "High income: OECD" and "High income:
nonOECD" groups?  


```{r }


load("../gdp_edu.Rda")

sort_gdp_edu <- gdp_edu[order(gdp_edu$economy_dollars),] 

high_income_OECD <- gdp_edu[gdp_edu$Income.Group == "High income: OECD",]
mean_oecd <-mean(high_income_OECD$rank)

high_income_nonOECD <- gdp_edu[gdp_edu$Income.Group == "High income: nonOECD",]
mean_nonoecd <-mean(high_income_nonOECD$rank)

```


=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  
__Distributional plots__  
-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  

4 Show the distribution of GDP value for all the countries and color plots by income group. Use
ggplot2 to create your plot.

```{r distribution_plots, echo = FALSE, message = FALSE}


ggplot(gdp_edu, aes(x = country_code, y = log10(economy_dollars), shape = Income.Group, color = Income.Group )) + geom_point()

ggplot(gdp_edu, aes(x = Income.Group, y = log10(economy_dollars), fill = Income.Group)) + geom_boxplot() +
    guides(fill=FALSE)

ggplot(gdp_edu, aes(x = log10(economy_dollars), fill =  Income.Group)) + geom_density(alpha=.3)

```


# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-
# ...	summary statistics
# ...	-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-

5 Provide summary statistics of GDP by income groups.
6 Cut the GDP ranking into 5 separate quantile groups. Make a table versus Income.Group.
How many countries are Lower middle income but among the 38 nations with highest
GDP?

```{r sort_and_group, echo = FALSE, message = FALSE}

library('dplyr')

xxx <- gdp_edu %>% group_by(Income.Group) %>% summarize(
				n_pays = length(Income.Group),
				min_gdp = min(economy_dollars),
				qtr_01 = quantile(economy_dollars)[[2]],
				mean_gdp = mean(economy_dollars),
				qtr_03 = quantile(economy_dollars)[[4]],
				max_gdp = max(economy_dollars),
				stdev = sd(economy_dollars))
				
sort_xxx <- xxx[order(xxx$mean_gdp),] 
				
sort_xxx

```

=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  
__Quartile Summary__  
-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  

```{r quintiles}

sort_gdp_edu_by_rank <- gdp_edu[order(gdp_edu$rank),] 

n_ranks <- dim(gdp_edu)[1]
quint <- c(0 : (n_ranks-1))
quint <- as.integer(quint/(n_ranks/5) ) + 1

sort_gdp_edu_by_rank <- cbind (sort_gdp_edu_by_rank, quint)

hgh_rank_lower_middle_income <- sort_gdp_edu_by_rank[
										sort_gdp_edu_by_rank$Income.Group == "Lower middle income"
										& quint == 1,]

sgebr <- sort_gdp_edu_by_rank


summary_table <- table(sgebr$quint, sgebr$Income.Group)
summary_table

barplot(summary_table,legend = T, beside = T, main = 'income by quintile')

library(vcd)

mosaicplot(summary_table, color = TRUE)

```
