---
title: "World Bank - Income Distributions and GDP"
author: "mcdevitt"
date: "8 mars 2017"
output:
  md_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(repmis)
library(RCurl)
library(methods)
library(ggplot2)
library(pastecs)
library(vcd)
require (reshape2)
require (xtable)
require (ascii)

home_dir <- "~/_smu/msds 6306 - data science/case_study_1/gdp_worldbank"
data_dir <- paste0(home_dir, "/data")
analysis_dir <- paste0(home_dir, "/analysis")
summary_dir <- paste0(home_dir, "/summary")
```

```{r }
# ...	load the saved data frame ...
load("../data/gdp_edu.Rda")
```


## Analysis Requested

### 1 : Merge the data based on the country shortcode. How many of the IDs match?  

####  Number of countries with matched records between the GDP and Educational data sets :  `r dim(gdp_edu)[1]`  

### 2 : Sort the data frame in ascending order by GDP (so United States is last). What is the 13th country in the resulting data frame?  
```{r}

sort_gdp_edu <- gdp_edu[order(gdp_edu$economy_dollars),] 

head(sort_gdp_edu[,1:4], 13)

```

#### The 13th ranked nation by GDP is : __`r sort_gdp_edu$country_name[13]`__  


### 3 : What are the average GDP rankings for the "High income: OECD" and "High income:
nonOECD" groups?  

```{r }

high_income_OECD <- gdp_edu[gdp_edu$Income.Group == "1_High_OECD",]

high_income_nonOECD <- gdp_edu[gdp_edu$Income.Group == "2_High_nonOECD",]

```

#### The mean ranking for HighIncome:OECD    nations is :  __`r mean(high_income_OECD$rank)`__  
#### The mean ranking for HighIncome:nonOECD nations is : __`r mean(high_income_nonOECD$rank)`__  

### __Distributional plots__  

### 4 Show the distribution of GDP value for all the countries and color plots by income group.

```{r distribution_plots, message = FALSE}

ggplot(gdp_edu, aes(x = country_code, y = log10(economy_dollars), shape = Income.Group, color = Income.Group )) + geom_point()

ggplot(gdp_edu, aes(x = Income.Group, y = log10(economy_dollars), fill = Income.Group)) + geom_boxplot() +
    guides(fill=FALSE)

ggplot(gdp_edu, aes(x = log10(economy_dollars), fill =  Income.Group)) + geom_density(alpha=.3)

```


## __... Summary Statistics__  

### 5 Provide summary statistics of GDP by income groups.  

```{r sort_and_group, message = FALSE}

library('dplyr')

gdp_edu_table <- gdp_edu %>% group_by(Income.Group) %>% summarize(
				n_nation = length(Income.Group),
				min_gdp = min(economy_dollars),
				qtr_01 = quantile(economy_dollars)[[2]],
				mean_gdp = mean(economy_dollars),
				qtr_03 = quantile(economy_dollars)[[4]],
				max_gdp = max(economy_dollars),
				stdev = sd(economy_dollars))
				
sort_gdp_edu_table <- gdp_edu_table[order(gdp_edu_table$mean_gdp),] 

#### Summary statistics for the 5 income groups are shown in following table.  
#The summary statistics include :  
#- Number of countries in each group  
#- Minimum GDP in each group (Millions $ US)  
#- 25th percentile GDP  
#- Mean GDP  
#- 75th percentile GDP  
#- Maximum GDP in the group  

sort_gdp_edu_table

```



=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-  
__Quartile Summary__  
-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-   
### 6 Cut the GDP ranking into 5 separate quantile groups.  
Make a table versus Income.Group.  

```{r quintiles}

# ...	to identify the quintile gdp ranking ... create new data frame, sorted by rank
sort_gdp_edu_by_rank <- gdp_edu[order(gdp_edu$rank),] 

# ...	count the number of ranks in the data set
n_ranks <- dim(gdp_edu)[1]

# ...	assign ranked order quintile (1..5) to each data row
quint <- c(0 : (n_ranks-1))
quint <- as.integer(quint/(n_ranks/5) ) + 1

# ...	bind the quintile group as additional column in data frame
sort_gdp_edu_by_rank <- cbind (sort_gdp_edu_by_rank, quint)

high_rank_lower_middle_income <- sort_gdp_edu_by_rank[
										sort_gdp_edu_by_rank$Income.Group == "4_Lower_Middle"
										& quint == 1,]
```

#### The list of countries which are Lower middle income but among the 38 nations with highest GDP  are :  

```{r}

print(sprintf(" N  : Rank  |            Country Name        | GDP (US$M)"))
for (indx in 1 : dim(high_rank_lower_middle_income)[1])
{
	print(sprintf("%2d : %5d | %30s | %8d", indx,
				  high_rank_lower_middle_income$rank[indx],
				  high_rank_lower_middle_income$Long.Name[indx],
				  high_rank_lower_middle_income$economy_dollars[indx]))
}

```

#### The following figures provide some visualizations of the distribution of countries, depicting the relationship between the quintiles of GDP distribition to income levels as identified from the World Bank Educational data sets. There is, as has been demonstrated above, a strong association between national GDP and per capita income level groupings.  

```{r}
# ...	shorten the data frame name, for convenience
sgebr <- sort_gdp_edu_by_rank

# ...	create summary table of quintile, income groups
summary_table <- table(sgebr$quint, sgebr$Income.Group)
summary_table

# ...	make some plots to visualize the data by quintile groups

barplot(summary_table,legend = T, beside = T, main = 'income by quintile')
mosaicplot(summary_table, color = TRUE)

```
